---
title: "Statistics Project 2"
author: "David Rodden, Michael McIntosh, Khang Tran, Jeffrey Lazatin"
date: "April 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(mosaic); require(mosaicData); require(gridExtra); require(knitr)
library("mosaic", lib.loc="~/R/win-library/3.3")
library("gridExtra", lib.loc="~/R/win-library/3.3")
library(readxl)
```
```{r}
sheet = read_excel("GDP.xlsx")
attach(sheet)
#Country, GDP, LEB, NLLEB, NLGDP
```

*Introduction*

We are investigating the relationship between Life Expectancy of a country based upon its GDP. The data we are using was collected in 2003 from the CIA Factbook; the data is across 180 countries. The investigation is looking to see if there is a positive correlation between life expectancy (LEB) and a country's GDP (NLGDP); using GDP as a predictor. In order to normalize the data, we use the natural log of the GDP. The data considers a country's life expectancy at birth and the GDP per capita (PPP). The data was collected from official reports that each nation compiles. We found the data from *Index Mundi*, who pulled from the CIA Factbook. The data is a sample of the world's countries, and is an observational study.


$H_0 : \rho = 0$ vs $H_a : \rho \neq 0$


*Summary & Visualization*

```{r}
favs = favstats(LEB ~ NLGDP)
anova.b = anova(lm(LEB ~ NLGDP))
xyplot(LEB ~ NLGDP, type = c("p", "r"), pch=16, col="black")
```
The scatter plot shows that there are a few outliers which will influence the overall model. The outliers will impact the regression which we use to model and predict, based upon the data. There is a slight departure from linearity, a subtle curve in the data, but still increasing overall. The data does possesses changing variability, a fanning trend, wide to narrow from left to right. There appears to be a positive linear association between the two quantitative variables, LEB ~ NLGDP. 


```{r}
#five number summary
sum.sheet = summary(sheet); sum.sheet
#standard deviation
gdp.sd = sd(sheet$GDP); gdp.sd
leb.sd = sd(sheet$LEB); leb.sd
nlgdp.sd = sd(sheet$NLGDP); nlgdp.sd
```
The sample size is 180 countries. The means for GDP, Life expectancy at birth, and Natural log of GDP are `r sum.sheet[4,2]`, `r sum.sheet[4,3]`, `r sum.sheet[4,4]`, respectively.

The standard deviation for GDP, Life expectancy, at birth and Natural log of GDP are $`r gdp.sd`$, $`r leb.sd`$, $`r nlgdp.sd`$ respectively.

*Correlation Test*

$H_0 : \rho = 0$ vs $H_a : \rho \neq 0$

```{r}
corr = cor.test(NLGDP, LEB)
corr.p.value = corr$p.value
```
The p-value $\approx `r corr.p.value`$.
As the p-value is very small, we reject the null hypothesis in favor of the alternative hypothesis that there is a non-zero correlation between Life expectancy at birth and the natural log of GDP.

*Regression*

```{r}
mod = lm(LEB ~ NLGDP)
y.hat = makeFun(mod); y.hat
mod.prediction = predict(mod); head(mod.prediction)
mod.resid = resid(mod); head(mod.resid)
mod.coef = unname(coef(mod)); mod.coef
```
Using the regression model, $\hat{Y} = \hat{\beta_0} + \hat{\beta_1}x = `r mod.coef[1]` + `r mod.coef[2]`x$
We predict a value of $\hat{y} \approx `r mod.prediction[1]`$ using x $\approx `r NLGDP[1]`$ with a residual $\approx `r mod.resid[1]`$.

```{r}
sd.verify = max(na.omit(favs$sd)) > 2 * min(na.omit(favs$sd)); sd.verify
n.verify = max(na.omit(favs$n)) >= 30; n.verify
```
Verifying conditions, we see that our data fails the standard deviation check for ANOVA.

Despite the conditions for ANOVA testing not meeting their requirements, we will proceed with the ANOVA test anyhow.
```{r}
anova(mod)
r = cor(LEB ~ NLGDP); r
n = length(NLGDP)
t = r * sqrt((n - 2) / (1 - r ^ 2)); t
p.value = 2 * pt(-abs(t), df = n - 2)
```
Our evidence allows us to reject the null hypothesis in favor of the alternative hypothesis. We have a t-statistic value .
Given the t statistic, $\approx `r t`$, and the p-value, $\approx `r p.value`$, we reject the null hypothesis with evidence for the alternative hypothesis. Our evidence from the data collected implies that there is a correlation between a country's life expectancy at birth and GDP.

*Teamwork*

As a team, we worked on coordinating meeting times. During the meetings, Michael collected and recorded data. After research, we discussed and analyzed it as a team. Following this, Michael and David typed the code within the RStudio environment as Jeffrey and Khang debugged and gave feedback. All data was protected by David Rodden's ingenious decision to upload resources to GitHub (https://github.com/DavidRodden/Statistics_Project_B). If any questions were had, Michael would follow up with the instructor for clarification.


```{r,include=FALSE}
kable(sheet)
```