---
title: "Statistics Project 2"
author: "David Rodden, Michael McIntosh, Khang Tran, Jeffrey Lazatin"
date: "April 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(mosaic); require(mosaicData); require(gridExtra); require(knitr)
library("mosaic", lib.loc="~/R/win-library/3.3")
library("gridExtra", lib.loc="~/R/win-library/3.3")
library(readxl)
```
```{r}
sheet = read_excel("GDP.xlsx")
attach(sheet)
#Country, GDP, LEB, NLLEB, NLGDP
```

*Introduction*

We are investigating the relationship between Life Expectancy of a country based upon its GDP. The data we are using was collected in 2003 from the CIA Factbook; the data is across 180 countries. The investigation is looking to see if there is a positive correlation between life expectancy (LEB) and a country's GDP (NLGDP); using GDP as a predictor. In order to normalize the data, we use the natural log of the GDP. The data considers a country's life expectancy at birth and the GDP per capita (PPP). The data was collected from official reports that each nation compiles. We found the data from *Index Mundi*, who pulled from the CIA Factbook. The data is a sample of the world's countries, and is an observational study.


$H_0 : \rho = 0$ vs $H_a : \rho \neq 0$


*Summary & Visualization*

```{r}
favs = favstats(LEB ~ NLGDP)
anova.b = anova(lm(LEB ~ NLGDP))
xyplot(LEB ~ NLGDP, type = c("p", "r"), pch=16, col="black")
```
The scatter plot shows that there are a few outliers which will influence the overall model. The outliers will impact the regression which we use to model and predict, based upon the data. There is a slight departure from linearity, a subtle curve in the data, but still increasing overall. The data does possesses changing variability, a fanning trend, wide to narrow from left to right. There appears to be a positive linear association between the two quantitative variables, LEB ~ NLGDP. 


```{r}
#five number summary
sum.sheet <- summary(sheet)
sum.sheet
#standard deviation
leb.sd <- sd(sheet$LEB)
nlgdp.sd <- sd(sheet$NLGDP)
```
The sample size is 180 countries. The means for GDP, Life expectancy at birth, and Natural log GDP are `r sum.sheet[4,2]`, `r sum.sheet[4,3]`, `r sum.sheet[4,4]`, respectively.


*Correlation Test*


*Regression*


*Teamwork*



```{r,include=FALSE}
kable(sheet)
```